Embedding Dimension of Polyhedral Losses

### 2020-01-30 colt-20/embed-dim.tex
Fixing some todos in proofs

\begin{theorem} \label{thm:general-char}
  Let $\ell: \R \to \reals_{\geq 0}^{\Y}$ be a discrete loss with, for each $r \in \R$, $\gamma_r = \{p : r \in \argmin_{r'} \inprod{p}{\ell(r)}\}$.
  Then $\ell$ is $d$-embeddable if and only if there exist polytopes $\T = \{T_y \subseteq \reals^d : y \in \Y\}$ for each $r \in \R$ such that both of the following hold:
  \begin{enumerate}
    \item (Optimality) We have $D(\T) = \gamma_r$.
    \item (Monotonicity) There exists an injective embedding function $\varphi : \R \to \reals^d$ and loss functions $\{L_y : \reals^d \to \reals^\Y_{\geq 0}\}_{y \in \Y}$ such that for all $y \in \Y$, we have $T_y = \partial L_y(\varphi(r))$.
  \end{enumerate}
\end{theorem}
\begin{proof}
  \jessie{An attempt:}
  $(\implies)$ If $\ell$ is $d$-embeddable, say that $L:\reals^d \to \reals^\Y_{\geq 0}$ embeds $\ell$.
  Take $T_y$ to be $\partial L(\varphi(r))_y$ for a fixed $r \in \R$ and $y \in \Y$.
  Clearly this set of polytopes $\T$ satisfies the monotonicity condition.
  To see the optimality condition is satisfied, consider that $\gamma_r = \Gamma_{\varphi(r)} = \{p \in \simplex : \vec 0 \in \oplus_p \partial L(\varphi(r))_y \} = \{p : \vec 0 \in \oplus_p T_y \} = D(\T)$.

  \bigskip
  $(\impliedby)$ The first embedding condition holds by the assumption of $\varphi$ in the monotonicity condition.
  Now consider that $\gamma_r = D(\T) = \{p \in \simplex : \vec 0 \in \oplus_p T_y \} = \{p \in \simplex : 0 \in \oplus_p \partial L(\varphi(r))_y \} = \{p : r \in \Gamma(p) \} =  \Gamma_{\varphi(r)}$.
  Therefore, as this holds for all $r \in \R$, we have that the Bayes Risks of the two losses match, and by~\cite[Proposition 1]{finocchiaro2019embedding}, we have $L$ embedding $\ell$. 
  \jessiet{This is a hole, since the subgradients of the level sets match, but the values do not.  Need to bake matching losses into monotonicity condition}
\end{proof}

@@

To gain a better understanding about the QFP in Definition~\ref{def:qfp}, we prove the following:
\begin{lemma} \label{lemma:D-polytope}
  For any $\T$, $D(\T)$ is a polytope (in particular, is convex).
\end{lemma}
\begin{proof}
   Recall by definition, the notation $\oplus_p \T = \{\sum_y p(y) x^r_y : x^r_y \in T_y (\forall y)\}$.
   Each $T_y$ is a polytope, so $p(y) T_y$ is a polytope.
   The Minkowski sum of polytopes is a polytope, so $\oplus_p \T$ is a polytope by~\cite[Section 1.2]{weibel2007minkowski}.
  \bo{TODO}
\end{proof}
\begin{lemma}  \label{lemma:minkowski-support}
  Given polytopes $\T$, there exists a finite set of normal vectors $w_1,\ldots,w_K \in \reals^d$ such that, for all $p \in \simplex$, $\oplus_p \T = \{x : \inprod{w_i}{x} \leq \sum_{y \in \Y} p(y) \max_{x \in T_y} \inprod{w_i}{x} \}$.
\end{lemma}
\begin{proof}
   For each $p$, $\oplus_p \T$ is a polytope.
  \bo{TODO: it turns out that $p$ doesn't matter, only its support; and there are only finitely many possible supports (or even better, the normals that are complete for the uniform minkowski sum turn out to be complete for all other $p$-combinations).}
\end{proof}
\begin{lemma} \label{lemma:E-to-B}
  Let $C = \{p : Bp \geq \vec 0 \}$ where $B$ has the minimum possible number of rows to capture $C$, and suppose $C = \{p : Ep \geq \vec 0 \}$.
  Then for each row in $B$ there is some (unique) row in $E$ that is equal to $\alpha B$ for some positive $\alpha$.
\end{lemma}
\begin{proof}
  \bo{? Cite} \jessie{Is left as an exercise in the following text:} \cite[Exercise 2.15]{ziegler2012lectures}
\end{proof}

@@

\bo{Am assuming that we will have to move almost everything to the appendix, wrote this blurb instead.}\jessie{Looks good to me.}
We have two proofs, both delegated to the appendix.
The first is to directly obtain from mathematical software that the QFP is infeasible.
We use observations about $\ell_{1/2}$ to make a number of simplifications to the program first.
This reduces the QFP to the following problem: 
\begin{quotation}
  Given a real-valued matrix $M$ in which some entries are missing, but have bounds of the form $[a,b]$, do there exist legal values for these entries such that $M$ is rank $d$?
\end{quotation}
The second proof is a direct geometric one utilizing the same observations.
We show that packing too many normal vectors into a half-circle yields contradictions, implying that a $5$-outcome embedding into $2$ dimensions is impossible (and incidentally characterizing the structure of all possible $4$-dimensional embeddings).

>>>>

\begin{theorem} \label{thm:general-char}
  Let $\ell: \R \to \reals_{\geq 0}^{\Y}$ be a discrete loss with, for each $r \in \R$, $\gamma_r = \{p : r \in \argmin_{r'} \inprod{p}{\ell(r)}\}$.
  Then $\ell$ is $d$-embeddable if and only if there exist polytopes $\T = \{T_y \subseteq \reals^d : y \in \Y\}$ for each $r \in \R$ such that both of the following hold:
  \begin{enumerate}
    \item (Optimality) We have $D(\T) = \gamma_r$.
    \item (Monotonicity) There exists an injective embedding function $\varphi : \R \to \reals^d$ and loss functions $\{L_y : \reals^d \to \reals^\Y_{\geq 0}\}_{y \in \Y}$ such that for all $y \in \Y$, we have $T_y = \partial L_y(\varphi(r))$ and for all $r \in \R$, we have $L_y(\varphi(r)) = \ell_y(r)$.
  \end{enumerate}
\end{theorem}
\begin{proof}
  \jessie{An attempt:}
  $(\implies)$ If $\ell$ is $d$-embeddable, say that $L:\reals^d \to \reals^\Y_{\geq 0}$ embeds $\ell$.
  Take $T_y$ to be $\partial L(\varphi(r))_y$ for a fixed $r \in \R$ and $y \in \Y$.
  Clearly this set of polytopes $\T$ satisfies the monotonicity condition.
  To see the optimality condition is satisfied, consider that $\gamma_r = \Gamma_{\varphi(r)} = \{p \in \simplex : \vec 0 \in \oplus_p \partial L(\varphi(r))_y \} = \{p : \vec 0 \in \oplus_p T_y \} = D(\T)$.

  \bigskip
  $(\impliedby)$ The first embedding condition holds by the assumption of $\varphi$ in the monotonicity condition.
  Now consider that $\gamma_r = D(\T) = \{p \in \simplex : \vec 0 \in \oplus_p T_y \} = \{p \in \simplex : 0 \in \oplus_p \partial L(\varphi(r))_y \} = \{p : r \in \Gamma(p) \} =  \Gamma_{\varphi(r)}$.
  Therefore, as this holds for all $r \in \R$, and we have matching losses on the embedding points by monotonicity, we have that the Bayes Risks of the two losses match, and by~\cite[Proposition 1]{finocchiaro2019embedding}, we have $L$ embedding $\ell$. 
\end{proof}

@@

To gain a better understanding about the QFP in Definition~\ref{def:qfp}, we prove the following:
\begin{lemma} \label{lemma:D-polytope}
  For any $\T$, $D(\T)$ is a polytope (in particular, is convex).
\end{lemma}
\begin{proof}
  \jessie{Appendix}
   Recall by definition, the notation $\oplus_p \T = \{\sum_y p(y) x_y : x_y \in T_y (\forall y)\}$.
   Each $T_y$ is a polytope, so $p(y) T_y$ is a polytope.
   The Minkowski sum of polytopes is a polytope, so $\oplus_p \T$ is a polytope by~\cite[Section 1.2]{weibel2007minkowski}.
  \bo{TODO}\jessie{Taking a try}
  Since $\oplus_p \T$ is a polytope for all $p \in \simplex$, we know there is a halfspace representation of normals $V$ so that for all $y \in \Y$, we have $x \in p(y)T_y \iff \inprod{V}{x} \leq p(y) e^y$ for some matrix $V$ and the support vector $e^y$, where $e^y_i = H_{T_y}(V_i)$.
  Adding redundant constraints for the $e^y$ as necessary, we know that the support of the Minkowski sum for a given normal is the sum of the normals~\cite[Theorem 3.1.6]{weibel2007minkowski}.
  Concatenating the vectors $e^y$ into $E$, we have $x \in \oplus_p \T \iff \inprod{V^*}{x} \leq E^*p$, where $V^*$ has redundant rows added as necessary. \jessie{Hand wavy... not a fan.}
  Substituting $x = \vec 0$, we see $\vec 0 \in \oplus_p \T \iff Ep \geq \vec 0$, which defines a polytope by construction of $E$.
\end{proof}
\begin{lemma}  \label{lemma:minkowski-support}
  Given polytopes $\T$, there exists a finite set of normal vectors $w_1,\ldots,w_K \in \reals^d$ such that, for all $p \in \simplex$, $\oplus_p \T = \{x : \inprod{w_i}{x} \leq \sum_{y \in \Y} p(y) \max_{x \in T_y} \inprod{w_i}{x} \}$.
\end{lemma}
\begin{proof}
  \jessie{Appendix}
   For each $p$, $\oplus_p \T$ is a polytope.
  \bo{TODO: it turns out that $p$ doesn't matter, only its support; and there are only finitely many possible supports (or even better, the normals that are complete for the uniform minkowski sum turn out to be complete for all other $p$-combinations).}
  For each of the finitely many supports ($2^n - 1$), we know $\oplus_p \T$ is a polytope, and every polytope can be uniquely defined by an enumeration of vectors normal to its facets.
  As a two polytopes with the same support can be defined by the same facet enumeration, we can simply concatenate these finite set of normals for the finite polytope supports.
  Vectors that are not normal to a facet simply become redundant constraints.
  This yields finitely many normals defining the weighted Minkowski sum $\oplus_p \T$ for all $p \in \simplex$.
\end{proof}
\begin{lemma} \label{lemma:E-to-B}
  Let $C = \{p : Bp \geq \vec 0 \}$ where $B$ has the minimum possible number of rows to capture $C$, and suppose $C = \{p : Ep \geq \vec 0 \}$.
  Then for each row in $B$ there is some (unique) row in $E$ that is equal to $\alpha B$ for some positive $\alpha$.
\end{lemma}
\begin{proof}
  \bo{? Cite} \jessie{Is left as an exercise in the following text:} \cite[Exercise 2.15]{ziegler2012lectures}
\end{proof}

@@

\bo{Am assuming that we will have to move almost everything to the appendix, wrote this blurb instead.}\jessie{Looks good to me.}
We have two proofs, both delegated to the appendix.
The first is to directly obtain from mathematical software that the QFP is infeasible.
We use observations about $\ell_{1/2}$ to make a number of simplifications to the program first.
This reduces the QFP to the following problem: 
\begin{quotation}
  Given a real-valued matrix $M$ in which some entries are missing, but have bounds of the form $[a,b]$, do there exist legal values for these entries such that $M$ is rank $d$?
\end{quotation}
The second proof is a direct geometric one utilizing the same observations.
We show that packing too many normal vectors into a half-circle yields contradictions, implying that a $5$-outcome embedding into $2$ dimensions is impossible (and incidentally characterizing the structure of all possible $4$-outcome\jessie{dimensional $\to$ outcome} embeddings).
